Alleviating the Topology Mismatch Problem in Distributed Overlay Networks: A Survey

### 2015-09-13 background.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

{\sl Napster} was the most successful incarnation of the
centralized approach in file-sharing~\cite{napster}.
It maintained a \emph{central index server} based on file-lists 
that were peer-provided.
Users would query the index server that would provide 
pointers to the actual content.
Thus, by centralizing search while distributing downloads,  {\sl Napster}
achieved a highly efficient design that, at its height, attracted $26$ million
users sharing approximately $80$ million songs\cite{jmm_naptopusage_2001}.
%%
The central indexing component played a vital role in the demise of 
{\sl Napter}, as the \emph{Recording
Industry Association of America} successfully argued in court that this software
entity was enabling copyright infringement of the artists' songs exchanged
through the system.

Architectures that distributed both search and resource provision 
capabilities (Figure~\ref{figure:p2p-archs:unstructred-full}) soon followed.
In these \emph{decentralized} approaches, 
resource placement within the overlay topology is random~\cite{YG-M2002};
for this reason, such architectures are referred to as \emph{unstructured}. 
Key properties of such \p\ systems are that they support the inherent heterogeneity
of peers, are highly resilient to peer failures, and incur low maintenance
overhead while handling the dynamics of peer participation
\cite{stutzbach_churn_2006}. 
Unstructured systems are also known as 
\emph{broadcast-based} systems for they use message flooding among %% it is FOR!
peers to propagate search queries. 
In its first realization,
{\sl Gnutella}(\emph{v.0.4}) became a well-known example of the fully
decentralized unstructured approach.

A key disadvantage of the unstructured approach is that message flooding burdens
the network since queries travel within the network randomly, visiting nodes
that do not have the sought resource and thus wasting bandwidth.  
To save on bandwidth, unstructured networks typically limit how long 
a query travels within the overlay via the \emph{time-to-live} parameter
(\emph{TTL}). 
When a query is received by a node, it decreases its \emph{TTL} by one unit. 
If the \emph{TTL} has not reached zero, the node forwards the query 
to its neighbors.  Otherwise, it drops the query. 
While this approach prevents queries from visiting the entire network, it
does not guarantee that the resource of interest, (e.g., a rare recording)
will, ultimately, be found.
%%%
\begin{figure}[ht]
\centering
\subfigure[Centralized ({\sl SETI@Home} or {\sl Napster}). Peers query the centralized
entity (dotted lines) and then establish the connection to the data source.] {
  \includegraphics[scale=0.17]{img/pdf/centralized.pdf}
  \label{figure:p2p-archs:centralized}
}\qquad\qquad
\subfigure[Fully unstructured ({\sl Gnutella} \emph{v.0.4}). Data is distributed randomly,
the topology is also random and all participants are equal.] {
  \includegraphics[scale=0.17]{img/pdf/unstructured-full.pdf}
  \label{figure:p2p-archs:unstructred-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured ({\sl Gnutella} \emph{v.0.6} or {\sl FastTrack}). Certain
peers (bigger circles) shoulder specific responsibilities (e.g., data
indexing).] {
  \includegraphics[scale=0.16]{img/pdf/unstructured-hierarchical.pdf}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad%\qquad
\subfigure[Structured ({\sl Chord}~\cite{stoica_chord_2001} and {\sl Kademlia}~\cite{maymounkov_kademlia_2002}). 
Peers and data are placed in the overlay in a deterministic manner.] {
  \includegraphics[scale=0.17]{img/pdf/structured.pdf}
  \label{figure:p2p-archs:structured}
}
\caption{The various \p\ architectures.}
\label{figure:p2p-archs}
\end{figure}

@@

To shorten search times, improve scalability, and lighten network load, 
unstructured configurations evolved into \emph{hierarchical} systems 
that dynamically assigned indexing functions to special peers, 
called \emph{ultra-}  or \emph{super-peers} 
(Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). 
Peers with reliable network connections and/or high compute power 
became \emph{ultra/super-peers}
facilitating less-powerful \emph{leaf-nodes} to find resources of
interest. 
Examples of such hierarchical unstructured systems include
{\sl Gnutella}(\emph{v.0.6})/{\sl Limewire}~\cite{gnutella}, 
{\sl KaZaA}~\cite{kazaa} and {\sl Skype}~\cite{skype}. 
Although the hierarchical approach relieves network pressure as
search queries were flooded over the much smaller subset of 
ultra/super-peers, it fails to address the reduced search scope and
inability to locate rare resources.

\emph{Decentralized structured schemes} were introduced as an answer to the above
problems of decentralized unstructured \p--approaches
(Figure~\ref{figure:p2p-archs:structured}).
Here, the main objective is to offer 
a self-organizing infrastructure for large-scale \p\ applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
To achieve this, they implement a \emph{Distributed Hash Table (DHT)} that maps
objects to nodes through a deterministic mechanism. \emph{DHT}s provide a
guaranteed bound on the number of overlay routing hops that have to be
``traveled'' to locate a resource, even in the case when only a \emph{single}
copy exists within the system. This requires
$O$($log(n)$) hops, compared to the unstructured networks that require
$O(n)$ to reliably locate any object. Unfortunately this efficiency does not
come without cost. Structured overlays can only support exact-match queries 
(i.e., queries that identify resources by name) as opposed
to content-based retrieval provided by unstructured overlays. 
This means that unstructured overlays can support more versatile 
resource location queries such as keyword and partial matching searches. 
Several popular file-sharing \p\ applications have been based on \emph{DHT}s
including {\sl Overnet}/{\sl eDonkey}~\cite{overnet}, 
{\sl Kademlia}/{\sl Kad Network}~\cite{maymounkov_kademlia_2002},  and 
some flavors of {\sl BitTorrent}~\cite{c_bittorrent_2003}.  
Moreover, \emph{DHT}s sparked
a flurry of research for many diverse applications, ranging
from distributed search engines to event-notification and cloud-based
platforms~\cite{kbc_oceanstore_2000,rkcd_scribe_2001,mgpj_cloudsnap_2011}.

@@

Early overlay protocol implementations gave litte attention to the 
topology mismatch problem.
{\sl Gnutella}, for example, is
considered far from scalable~\cite{ritter_gnucantscale_2001} for every peer
chooses its neighbors without any knowledge of the underlying
network, resulting in mismatches. 
Queries are flooded over multiple paths in an effort 
to reach nodes that have the sought content. 
In addition, 
pairs of nodes may simultaneously forward the same query to each other
in a superfluous mode of operation in 
this opportunistic and without coordination environment.
%%
\cite{matei_mapgnutella_2002} showed that, even when $95$\% of any two nodes 
are less than $7$ hops away from each other, 
a \emph{flooding-based} {\sl Gnutella}-like algorithm 
can generate $330$TB/month traffic in a $50,000$ node network. 
A topology mismatch can thus have a serious effect 
on Internet's routing performance.

@@

We employ both tabular and pictorial means to 
articulate features of the proposed methodologies and outline
both strengths and weaknesses.
To ascertain the effectiveness of each proposal, we predominantly use 
three criteria that capture the behavior of the algorithms 
discussed:
\emph{efficiency}, \emph{overhead} and \emph{scalability}.
For each of these criteria, we use a coarse-grained,
three-level, gradation to denote how well the algorithms do,
namely, \emph{low}, \emph{medium} and \emph{high}.
Each criterion has its own
interpretation of the above literal values. 
\emph{Efficiency} means how well the
algorithm performs its search and topology adaptation functions. Our assessment
also considers any negative effects that might come along with the application
of the algorithm on intrinsic characteristics of the system (i.e., search scope
coverage, download performance etc.).

For the \emph{overhead} criterion, we evaluate the cost
of computing and communicating peer join, message forwarding, 
and overlay topology maintenance. For instance, the overhead that an algorithm,
which needs global knowledge of the overlay topology, imposes on the system is
greater compared to a protocol that only relies on peers exchanging routing
tables within a 2-hop radius from each other.

For the \emph{scalability} criterion, we consider system behavior with respect
to the number of participating peers it can reliably support. We characterize
protocols as being lowly, mediumly or highly scalable if they are proven to
effectively support hundreds, tenths of thousands, or millions of nodes,
respectively.

>>>>

\emph{Napster} was the most successful incarnation of the
centralized approach in file-sharing~\cite{napster}.
It maintained a \emph{central index server} based on file-lists 
that were peer-provided.
Users would query the index server that would provide 
pointers to the actual content.
Thus, by centralizing search while distributing downloads,  \emph{Napster}
achieved a highly efficient design that, at its height, attracted $26$ million
users sharing approximately $80$ million songs~\cite{jmm_naptopusage_2001}.
%%
The central indexing component played a vital role in the demise of 
\emph{Napter}, as the \emph{Recording
Industry Association of America} successfully argued in court that this software
entity was enabling copyright infringement of the artists' songs exchanged
through the system.

Architectures that distributed both search and resource provision 
capabilities (Figure~\ref{figure:p2p-archs:unstructred-full}) soon followed.
In these \emph{decentralized} approaches, 
resource placement within the overlay topology is random~\cite{YG-M2002};
for this reason, such architectures are referred to as \emph{unstructured}. 
Key properties of such \p\ systems are that they support the inherent heterogeneity
of peers, are highly resilient to peer failures, and incur low maintenance
overhead while handling the dynamics of peer participation
\cite{stutzbach_churn_2006}. 
Unstructured systems are also known as 
\emph{broadcast-based} systems for they use message flooding among %% it is FOR!
peers to propagate search queries. 
In its first realization,
\emph{Gnutella}(\emph{v.0.4}) became a well-known example of the fully
decentralized unstructured approach.

A key disadvantage of the unstructured approach is that message flooding burdens
the network since queries travel within the network randomly, visiting nodes
that do not have the sought resource and thus wasting bandwidth.  
To save on bandwidth, unstructured networks typically limit how long 
a query travels within the overlay via the \emph{time-to-live} parameter
(\emph{TTL}). 
When a query is received by a node, it decreases its \emph{TTL} by one unit. 
If the \emph{TTL} has not reached zero, the node forwards the query 
to its neighbors.  Otherwise, it drops the query. 
While this approach prevents queries from visiting the entire network, it
does not guarantee that the resource of interest, (e.g., a rare recording)
will, ultimately, be found.
%%%
\begin{figure}[ht]
\centering
\subfigure[Centralized (\emph{SETI@Home} or \emph{Napster}). Peers query the centralized
entity (dotted lines) and then establish the connection to the data source.] {
  \includegraphics[scale=0.17]{img/pdf/centralized.pdf}
  \label{figure:p2p-archs:centralized}
}\qquad\qquad
\subfigure[Fully unstructured (\emph{Gnutella} \emph{v.0.4}). Data is distributed randomly,
the topology is also random and all participants are equal.] {
  \includegraphics[scale=0.17]{img/pdf/unstructured-full.pdf}
  \label{figure:p2p-archs:unstructred-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured (\emph{Gnutella} \emph{v.0.6} or \emph{FastTrack}). Certain
peers (bigger circles) shoulder specific responsibilities (e.g., data
indexing).] {
  \includegraphics[scale=0.16]{img/pdf/unstructured-hierarchical.pdf}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad%\qquad
\subfigure[Structured (\emph{Chord}~\cite{stoica_chord_2001} and \emph{Kademlia}~\cite{maymounkov_kademlia_2002}). 
Peers and data are placed in the overlay in a deterministic manner.] {
  \includegraphics[scale=0.17]{img/pdf/structured.pdf}
  \label{figure:p2p-archs:structured}
}
\caption{The various \p\ architectures.}
\label{figure:p2p-archs}
\end{figure}

@@

To shorten search times, improve scalability, and lighten network load, 
unstructured configurations evolved into \emph{hierarchical} systems 
that dynamically assigned indexing functions to special peers, 
called \emph{ultra-}  or \emph{super-peers} 
(Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). 
Peers with reliable network connections and/or high compute power 
became \emph{ultra/super-peers}
facilitating less-powerful \emph{leaf-nodes} to find resources of
interest. 
Examples of such hierarchical unstructured systems include
\emph{Gnutella}(\emph{v.0.6})/\emph{Limewire}~\cite{gnutella}, 
\emph{KaZaA}~\cite{kazaa} and \emph{Skype}~\cite{skype}. 
Although the hierarchical approach relieves network pressure as
search queries were flooded over the much smaller subset of 
ultra/super-peers, it fails to address the reduced search scope and
inability to locate rare resources.

\emph{Decentralized structured schemes} were introduced as an answer to the above
problems of decentralized unstructured \p--approaches
(Figure~\ref{figure:p2p-archs:structured}).
Here, the main objective is to offer 
a self-organizing infrastructure for large-scale \p\ applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
To achieve this, they implement a \emph{Distributed Hash Table (DHT)} that maps
objects to nodes through a deterministic mechanism. \emph{DHT}s provide a
guaranteed bound on the number of overlay routing hops that have to be
``traveled'' to locate a resource, even in the case when only a \emph{single}
copy exists within the system. This requires
$O$($log(n)$) hops, compared to the unstructured networks that require
$O(n)$ to reliably locate any object. Unfortunately this efficiency does not
come without cost. Structured overlays can only support exact-match queries 
(i.e., queries that identify resources by name) as opposed
to content-based retrieval provided by unstructured overlays. 
This means that unstructured overlays can support more versatile 
resource location queries such as keyword and partial matching searches. 
Several popular file-sharing \p\ applications have been based on \emph{DHT}s
including \emph{Overnet}/\emph{eDonkey}~\cite{overnet}, 
\emph{Kademlia}/\emph{Kad Network}~\cite{maymounkov_kademlia_2002},  and 
some flavors of \emph{BitTorrent}~\cite{c_bittorrent_2003}.  
Moreover, \emph{DHT}s sparked
a flurry of research for many diverse applications, ranging
from distributed search engines to event-notification and cloud-based
platforms~\cite{kbc_oceanstore_2000,rkcd_scribe_2001,mgpj_cloudsnap_2011}.

@@

Early overlay protocol implementations gave litte attention to the 
topology mismatch problem.
\emph{Gnutella}, for example, is
considered far from scalable~\cite{ritter_gnucantscale_2001} for every peer
chooses its neighbors without any knowledge of the underlying
network, resulting in mismatches. 
Queries are flooded over multiple paths in an effort 
to reach nodes that have the sought content. 
In addition, 
pairs of nodes may simultaneously forward the same query to each other
in a superfluous mode of operation in 
this opportunistic and without coordination environment.
%%
\cite{matei_mapgnutella_2002} showed that, even when $95$\% of any two nodes 
are less than $7$ hops away from each other, 
a \emph{flooding-based} \emph{Gnutella}-like algorithm 
can generate $330$TB/month traffic in a $50,000$ node network. 
A topology mismatch can thus have a serious effect 
on Internet's routing performance.

@@

We employ both tabular and pictorial means to 
articulate features of the proposed methodologies and outline
both strengths and weaknesses.
To ascertain the effectiveness of each proposal, we predominantly use 
three criteria that capture the behavior of the algorithms 
discussed:
\emph{efficiency}, \emph{overhead} and \emph{scalability}.
For each of these criteria, we use a coarse-grained,
three-level, gradation to denote how well the algorithms do,
namely, \emph{low}, \emph{medium} and \emph{high}.
Each criterion has its own
interpretation of the above literal values.

We use the term \emph{efficiency} to gauge the competence of a given
algorithm to match the overlay to the underlying network. Minimization
of duplicate messages traveling the network, quick object location and
preservation of search scope as well as ability and
effectiveness to adapt in high churn environments are some of the
characteristics we consider when grading an approach on this criterion.

With \emph{overhead}, we, coarsely, quantify the additional cost a given approach
imposes to the underlying resources when applied to
\begin{inparaenum}[\itshape i\upshape)]
  \item create,
  \item adapt, and
  \item maintain
\end{inparaenum}
the overlay network while preserving its topology awareness. Computing and
communicating peer join and departure, message forwarding, updating routing tables
are some of things that we take into account to perform our assessement. Note
that this is the one criterion of the three we use that its literal values
``low'' and ``high'' must be interpreted as ``better'' and ``worse'' respectivelly.

For the \emph{scalability} criterion, we consider system behavior with respect
to the number of participating peers it can reliably support. We characterize
protocols as being lowly, mediumly or highly scalable if they are proven to
effectively support hundreds, tenths of thousands, or millions of nodes,
respectively. One important consideration here is the extent of self-organization
of a system's peers and thus the overall distributed nature of the overlay network.
The least a network relies on global knowledge to perform it functions the smoother
it scales.

### 2015-09-13 intro.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

The peer-to-peer paradigm gave impetus to two ``killer'' applications:
file-sharing and Internet telephony.
The {\sl Napster} file-sharing system was widely acknowledged as the 
``fastest growing Internet application ever'' in $2001$ when it topped
$26$ million users sharing over $80$ million songs.
Other file-sharing applications followed suit, 
including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
million concurrently-connected peers, as well as 
{\sl BitTorrent} connecting over $150$ million monthly users by January $2012$~\cite{bittorrentusers}.
\p\ telephony saw explosive growth with the advent of {\sl Skype}.
Since its introduction in $2003$,
{\sl Skype} has become extremely popular with more than $650$ million users 
in $2011$~\cite{skypetotalusers} and an astonishing 
$50$+ million concurrently-online users~\cite{skypesymusers}.
Moreover, it has consistently eroded  the traffic handled by 
traditional telephony carriers by slicing away a staggering 
$214$ billion minutes of pertinent traffic in $2013$ alone~\cite{skypetraffic}.

@@

For over a decade, researchers have extensively investigated  
various aspects of the topology mismatch problem.
This research area started to become active after  The mismatch problem is Interest around All interesting work around the topology We omitted centralized Centralized systems are not

The main objective of this paper is to offer a comprehensive 
survey of the work carried out in the area
and provide a
taxonomy of the proposed solutions. We point out synergies, as well as
similarities and differences in the published approaches. 
Ultimately, our goal is to help readers sift through 
the voluminous literature, to help them
understand the advantages and disadvantages of each work, and 
to provide them with enough perspective so that 
when the need arises, they are able to
select, amongst the different approaches, the one that is most suitable for
their particular application.
%%
The rest of the paper is organized as follows: 
In Section~\ref{section:background}, we provide background on
overlay architectures including centralized, decentralized-unstructured and
decentralized-structured \p\ systems.
We also formally define the problem of topology mismatch 
and offer the rationale behind our 
evaluation of the techniques discussed in this paper. 
Work in the area was inspired mostly by the volatile and fully
distributed nature of the decentrilized overlays so we skip
centralized systems and in Sections~\ref{section:unstructured} and~\ref{section:structured}
we outline the surveyed research efforts
for unstructured and structured \p\ overlays, respectively.
We conclude in Section~\ref{section:conclusion}.

>>>>

The peer-to-peer paradigm gave impetus to two ``killer'' applications:
file-sharing and Internet telephony.
The \emph{Napster} file-sharing system was widely acknowledged as the 
``fastest growing Internet application ever'' in $2001$ when it topped
$26$ million users sharing over $80$ million songs.
Other file-sharing applications followed suit, 
including \emph{Gnutella} and \emph{Limewire} enjoying $3$ 
million concurrently-connected peers, as well as 
\emph{BitTorrent} connecting over $150$ million monthly users by January $2012$~\cite{bittorrentusers}.
\p\ telephony saw explosive growth with the advent of \emph{Skype}.
Since its introduction in $2003$,
\emph{Skype} has become extremely popular with more than $650$ million users 
in $2011$~\cite{skypetotalusers} and an astonishing 
$50$+ million concurrently-online users~\cite{skypesymusers}.
Moreover, it has consistently eroded  the traffic handled by 
traditional telephony carriers by slicing away a staggering 
$214$ billion minutes of pertinent traffic in $2013$ alone~\cite{skypetraffic}.

@@

For over a decade, researchers have extensively investigated  
various aspects of the topology mismatch problem.
The main objective of this paper is to offer a comprehensive 
survey of the work carried out in the area
and provide a
taxonomy of the proposed solutions. We point out synergies, as well as
similarities and differences in the published approaches. 
Ultimately, our goal is to help readers sift through 
the voluminous literature, to help them
understand the advantages and disadvantages of each work, and 
to provide them with enough perspective so that 
when the need arises, they are able to
select, amongst the different approaches, the one that is most suitable for
their particular application.
%%
The rest of the paper is organized as follows: 
In Section~\ref{section:background}, we provide background on
overlay architectures including centralized, decentralized-unstructured and
decentralized-structured \p\ systems.
We also formally define the problem of topology mismatch 
and offer the rationale behind our 
evaluation of the techniques discussed in this paper. 
Work in the area was inspired mostly by the volatile and fully
distributed nature of the decentrilized overlays so we skip
centralized systems and in Sections~\ref{section:unstructured} and~\ref{section:structured}
we outline the surveyed research efforts
for unstructured and structured \p\ overlays, respectively.
We conclude in Section~\ref{section:conclusion}.
